---
title: "DQ Report - PJT 7: VHF"
output:
  html_document:
    df_print: paged
---


```{r include=FALSE}
### LIBRARIES ------------------------
# this script requiers dataquieR version 2.0.1
#packageVersion("dataquieR")
# if a different verison is installed use
#utils::install.packages(pkgs = "https://cran.r-project.org/src/contrib/Archive/dataquieR/dataquieR_2.0.1.tar.gz", repos = NULL)

library(knitr)
library(dataquieR)
library(data.table)
library(lubridate)
library(xfun)
library(pointblank)
library(digest)
library(ggplot2)

```

```{r include=FALSE}
### READ DATA -------------------

# # only the fallback if we start knitr directly from R Studio
if (!exists("OUTPUT_DIR_BASE")) OUTPUT_DIR_BASE <- dirname(dirname(getwd()))
if (!exists("PROJECT_NAME")) PROJECT_NAME <- "VHF"
if (!exists("OUTPUT_DIR_LOCAL")) OUTPUT_DIR_LOCAL <- paste0(OUTPUT_DIR_BASE, "/outputLocal/", PROJECT_NAME)
if (!exists("OUTPUT_DIR_GLOBAL")) OUTPUT_DIR_GLOBAL <- paste0(OUTPUT_DIR_BASE, "/outputGlobal/", PROJECT_NAME)
if (!exists("OUTPUT_DIR_TMP")) OUTPUT_DIR_TMP <- paste0(OUTPUT_DIR_LOCAL, "/tmp")
if (!exists("error_file")) error_file <- paste0(OUTPUT_DIR_BASE, "/outputLocal/", PROJECT_NAME, "/Errors/ErrorMessage.txt")
if (!exists("retrieve_file_cohort")) retrieve_file_cohort <- paste0(OUTPUT_DIR_BASE, "/outputLocal/", PROJECT_NAME, "/Cohort.csv")
if (!exists("retrieve_file_diagnoses")) retrieve_file_diagnoses <- paste0(OUTPUT_DIR_BASE, "/outputLocal/", PROJECT_NAME, "/Diagnoses.csv")

if(!file.exists(retrieve_file_cohort) || !file.exists(retrieve_file_diagnoses)){
  write("Ergebnisse nicht gefunden.", file = error_file)
  stop("Results not found - aborting.")
}

# read result csv
cohort_full <- read.csv(retrieve_file_cohort, sep = ";", colClasses = "character")
diagnoses_full <- read.csv(retrieve_file_diagnoses, sep = ";", colClasses = "character")

# read metadata
metadata_report <- read.csv("metadata.csv")
checks_report <- read.csv("contradictions.csv")
missings_report <- read.csv("missing-codes.csv")

# select columns supported with metadata from cohort
if(!all( metadata_report$VAR_NAMES %in% colnames(cohort_full) )){
  write("Metadaten passen nicht zur Kohorte.", file = error_file)
  stop("Metadata does not match Cohort - aborting.")
}
cohort_report <- cohort_full
cohort_report <- cohort_report[,metadata_report$VAR_NAMES]

# separate department DQA as its a list
cohort_departments <- cohort_full$encounter.departments

#extract known data types
datatyp_report <- as.data.table(t(metadata_report$DATA_TYPE))
colnames(datatyp_report) <- metadata_report$VAR_NAMES

#extract known value labels
value_labels <- as.data.table(t(metadata_report$VALUE_LABELS[2:7]))
colnames(value_labels) <- metadata_report$VAR_NAMES[2:7]
# function that returns only labels in string as list
get_labels <- function(string) {
  # regex to split at | num = : \s?\|?\s?\d\s\=\s
  # [-1] removes 1 = at start
  return(strsplit(string, "\\s?\\|?\\s?\\d\\s\\=\\s")[[1]][-1])
}
value_labels <- sapply(value_labels, get_labels)

#set up missing codes
missing_unknown <- missings_report$CODE_VALUE[2]
#item has integer/string type but cannot be converted to date for age calculation
missing_datetime <- missings_report$CODE_VALUE[3]
#item has correct type but is missing in value labels
missing_label <- missings_report$CODE_VALUE[4]


### Transform to dataquiR format -------------------

# function that sets integer values according to labels and missing codes
apply_labels <- function(vector, labels) {

  # unify casing, trim whitespace
  vector <- tolower(vector)
  labels <- tolower(labels)
  vector <- trimws(vector, which = "both")
  labels <- trimws(labels, which = "both")

  r <- NULL
  for(item in vector) {
    # default assumes value of item is missing in value labels
    value <- missing_label
    # if item is whitespace or NA set value to NA for correct dataquiR missingness
    if(item == "" || is.na(item) ) value <- NA
    else {
      i <- 0
      for(l in labels) {
        i <- i + 1
        # if item is found in labels set value to corresponding integer
        if(item == l) value <- i
      }
    }
    r <- c(r,value)
  }

  return(r)
}

#apply labels and fix typing per column - missmatched typing is returned as NA (NOT as missing code)
cohort_report$NTproBNP.valueQuantity.value <- as.double(cohort_report$NTproBNP.valueQuantity.value)
cohort_report$NTproBNP.code <- as.integer( apply_labels(cohort_report$NTproBNP.code, value_labels$NTproBNP.code) )
cohort_report$NTproBNP.codeSystem <- as.integer(  apply_labels(cohort_report$NTproBNP.codeSystem, value_labels$NTproBNP.codeSystem) )
cohort_report$NTproBNP.unit <- as.integer( apply_labels(cohort_report$NTproBNP.unit, value_labels$NTproBNP.unit) )
cohort_report$NTproBNP.unitSystem <- as.integer( apply_labels(cohort_report$NTproBNP.unitSystem, value_labels$NTproBNP.unitSystem) )
cohort_report$gender <- as.integer( apply_labels(cohort_report$gender, value_labels$gender) )
cohort_report$age <- as.integer(cohort_report$age)
cohort_report$encounter.start <- parse_date_time(cohort_report$encounter.start, "ymd HMS", truncated = 5)
cohort_report$encounter.end <- parse_date_time(cohort_report$encounter.end, "ymd HMS", truncated = 5)
cohort_report$NTproBNP.valueQuantity.comparator <- as.integer( apply_labels(cohort_report$NTproBNP.valueQuantity.comparator, value_labels$NTproBNP.valueQuantity.comparator) )

# verify correct typing
#sapply(cohort_report, class)

```

```{r include=FALSE}
### GENERATE REPORTS ----------------------

#function for logging errors inside knit output
knit_error <- function(f ,er_message) {
   tryCatch(
      { f },
        error = function(e){
            print(er_message)
            print(e)
        }
    )
}

knit_error( smith_dq_report <- dq_report(study_data = cohort_report,
                          meta_data  = metadata_report,
                          check_table = checks_report,
                          MISSING_LIST = missings_report,
                          label_col  = LABEL) ,
'Could not generate DQ report for Cohort. Error: ')

```

## Data Quality Report for the 7. Projectathon of MII: Use Case 'atrial fibrillation'

#### This report was created with the [dataquieR](https://dataquality.ship-med.uni-greifswald.de/tutorials2.html) (Version 2.0.1) Package for R.

# Information about Contradictions:

## Plot of Contradiction Checks:
#### 1001: The end date of an Encounter was before its start date. 
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error( smith_dq_report$long_format$con_contradictions$results[[1]]$SummaryPlot, 'Could not print Contradictions SummaryPlot for Cohort. Error: ')


```

## Contradiction Checks as Table:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error( smith_dq_report$long_format$con_contradictions$results[[1]]$SummaryTable, 'Could not print Contradictions SummaryTable for Cohort. Error: ')

```
# Information about Data Values:

## NTproBNP Distribution with Limit Checks:
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error({
plots <- list()
plots$ntprobnp <- con_limit_deviations(study_data = cohort_report, meta_data = metadata_report, resp_vars = 'NTproBNP.valueQuantity.value', limits = c("SOFT_LIMITS"))$SummaryPlotList$NTproBNP.valueQuantity.value
plots$ntprobnp$plot_env$p$coordinates$limits$x <- c(0,14500)
plots$ntprobnp
}, 'Could not print Limit NTproBNP Limit Plot for Cohort. Error: ')

```

## NTproBNP Limit Checks as Table:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(con_limit_deviations(study_data = cohort_report, meta_data = metadata_report, resp_vars = 'NTproBNP.valueQuantity.value')$SummaryTable, 'Could not print Limit SummaryPlot for Cohort. Error: ')

```

## NTproBNP Outlier Detection:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(acc_univariate_outlier(resp_vars = 'NTproBNP.valueQuantity.value',
                         study_data = cohort_report,
                         meta_data = metadata_report
                       )$SummaryPlotList$NTproBNP.valueQuantity.value, 'Could not print NTproBNP Outlier SummaryPlot for Cohort. Error: ')

```

## NTproBNP Outlier Detection as Table:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(acc_univariate_outlier(resp_vars = 'NTproBNP.valueQuantity.value',
                         study_data = cohort_report,
                         meta_data = metadata_report
                       )$SummaryTable, 'Could not print NTproBNP Outlier SummaryTable for Cohort. Error: ')

```



## Age Distribution with Limit Checks:
```{r message=FALSE, warning=FALSE,echo=FALSE}

#other possible limits: soft limit [18;120] hard limit [-1;200] ?

knit_error({
  plots$age <- con_limit_deviations(study_data = cohort_report, meta_data = metadata_report, resp_vars = 'age', limits = c("HARD_LIMITS"))$SummaryPlotList$age
  plots$age$plot_env$p$coordinates$limits$x <- c(0,125)
  plots$age
}, 'Could not print Age Limit Plot for Cohort. Error: ')

# deprecated age plot now in con_limit_deviations
# acc_plots in 2.0.1, dataquieR::report2()
# acc_plots <- acc_distributions(resp_vars      = NULL, 
#                          label_col      = "LABEL",
#                          study_data     = cohort_report, 
#                          meta_data      = metadata_report)
#knit_error(acc_plots$SummaryPlotList$age, 'Could not print SummaryPlotList$age for Cohort. Eorror: ')

```  

## Age Limit Checks as Table:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(con_limit_deviations(study_data = cohort_report, meta_data = metadata_report, resp_vars = 'age')$SummaryTable, 'Could not print Limit SummaryPlot for Cohort. Error: ')

```
## Age Outlier Detection:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(acc_univariate_outlier(resp_vars = 'age',
                         study_data = cohort_report,
                         meta_data = metadata_report
                       )$SummaryPlotList$age, 'Could not print Age Outlier SummaryPlot for Cohort. Error: ')

```

## Age Outlier Detection as Table:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(acc_univariate_outlier(resp_vars = 'age',
                         study_data = cohort_report,
                         meta_data = metadata_report
                       )$SummaryTable, 'Could not print Age Outlier SummaryTable for Cohort. Error: ')

```
# Information about Code Usage:

## Department Codes:
#### German Department Codes (Fachabteilungsschlüssel), see [code system](https://simplifier.net/packages/de.basisprofil.r4/1.4.0/files/656778) for full list.
#### 0100 = Innere Medizin | 0300 = Kardiologie | 0600 = Endokrinologie | 3700 = Sonstige Fachabteilung
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error( {
  cohort_departments_list <- unlist(strsplit(cohort_departments," ~ "))
  cohort_departments_table <- table( as.vector( cohort_departments_list) )
  tables <- list()
  tables$departments <- cohort_departments_table
} , 'Could not calculate information about Department Codes for Cohort. Error: ')

knit_error( {
  barplot( cohort_departments_table, horiz = FALSE, xlab = "Department Code", ylab = "Frequency" )
  }, 'Could not Plot Department Codes for Cohort. Error: ')
print('Department Codes as table:')
knit_error( cohort_departments_table, 'Could not print Department Code table for Cohort. Error: ')

knit_error( {
  print( paste0("Total columns: ", length(cohort_departments), "    Missing department codes (NA): ", 
  sum(is.na(cohort_departments)) ) )
  print( paste0("Total department codes: ", length(cohort_departments_list), "    Unique department codes: ",
  length(unique(cohort_departments_list))) )
} , 'Could not calculate information about Department Codes for Cohort. Error: ')

```

## Gender:
#### 1 = male | 2 = female | 3 = other | 4 = unknown
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error( 
tables$gender <-  table(cohort_report$gender), 'Could not create Table Gender. Error: ')

knit_error( #acc_plots$SummaryPlotList$gender
barplot( tables$gender, horiz = FALSE, xlab = "Gender", ylab = "Frequency" ), 'Could not print SummaryPlotList$gender for Cohort. Error: ')

print('Gender as table:')
knit_error(tables$gender, 'Could not tab SummaryPlotList$gender for Cohrot. Error: ')

```

## LOINC Codes:
#### 1 = other | 2 = 33763-4 | 3 = 71425-3 | 4 = 33762-6 | 5 = 83107-3 | 6 =  83108-1 | 7 = 77622-9 | 8 = 77621-1
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error( 
tables$loinc <-  table(cohort_report$NTproBNP.code), 'Could not create Table LOINC. Error: ')

knit_error(#acc_plots$SummaryPlotList$NTproBNP.code
  barplot( tables$loinc, horiz = FALSE, xlab = "LOINC Code", ylab = "Frequency" ), 'Could not print SummaryPlotList$NTproBNP.code for Cohrot. Error: ')

print('LOIC Codes as table:')
knit_error(tables$loinc, 'Could not tab SummaryPlotList$NTproBNP.code for Cohrot. Error: ')

```

## Comperators:
#### 1 = other | 2 = < | 3 = <= | 4 = > | 5 = >=
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error( 
tables$comperators <-  table(cohort_report$NTproBNP.valueQuantity.comparator), 'Could not create Table Comperators. Error: ')

knit_error(#acc_plots$SummaryPlotList$NTproBNP.valueQuantity.comparator
  barplot( tables$comperators, horiz = FALSE, xlab = "Comperator", ylab = "Frequency" ), 'Could not print SummaryPlotList$comparator for Cohort. Error: ')

print('Comperators as table:')
knit_error(tables$comperators, 'Could not tab SummaryPlotList$study_data$NTproBNP.valueQuantity.comparator for Cohrot. Error: ')

```

## Units:
#### 1 = other | 2 = pg/mL | 3 = pg/dL | 4 = pg/100mL | 5 = pg/L | 6 = pg% | 7 = pmol/L | 8 = ng/L | 9 = Mt/m3
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error( 
tables$units <-  table(cohort_report$NTproBNP.unit), 'Could not create Table Comperators. Error: ')

knit_error(#acc_plots$SummaryPlotList$NTproBNP.unit
  barplot( tables$units, horiz = FALSE, xlab = "Unit", ylab = "Frequency" ), 'Could not print SummaryPlotList$NTproBNP.unit for Cohort. Error: ')

print('Units as table:')
knit_error(tables$units, 'Could not tab SummaryPlotList$study_data$NTproBNP.unit for Cohrot. Error: ')

```

# Information about Completeness:

## Item Missingness as Table:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

#Table
knit_error(smith_dq_report$long_format$com_item_missingness$results[[1]]$SummaryTable, 'Could not print Missingness Table for Cohort. Error: ')

```

## Completeness with Missing Codes:
#### MISSING 9991100: Missing - reason unknown
#### MISSING 9991200: Deleted - Datetime format not supported
#### MISSING 9991300: Deleted - not in VALUE_LABELS
Regarding completeness and missing codes:
If a field has the wrong datatype it is evaluated to NA and tracked under Sysmiss in the completeness table. If a field has the correct typing but an undefined value (for example any string as a gender other than 'male', 'female', 'other', 'unknown') the missing code 9991300 is used instead.
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

#Plot
knit_error(#smith_dq_report$long_format$com_item_missingness$results[[1]]$ReportSummaryTable
  smith_dq_report$long_format$com_item_missingness$results[[1]]$SummaryPlot, 'Could not print Missingness SummaryPlot for Cohort. Error: ')

print("Missingness as table: ")
knit_error(#smith_dq_report$long_format$com_item_missingness$results[[1]]$ReportSummaryTable
  smith_dq_report$long_format$com_item_missingness$results[[1]]$ReportSummaryTable, 'Could not print Missingness SummaryPlot for Cohort. Error: ')


```

# Further possible Analysis with dataquieR:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(smith_dq_report$app_mat$ApplicabilityPlot, 'Could not print ApplicabilityPlot for Cohort. Error: ')

```

# Extract Plots for Cohort:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

 # make dir for plots
  dir.create(OUTPUT_DIR_TMP)
  TMP_DIR <- paste0(OUTPUT_DIR_TMP,"/plots")
  dir.create(TMP_DIR)

#function to embed variables as .rds in knitr output with xfun
embed_var = function(x, ...) {
  f <- tempfile(fileext = '.rds', tmpdir = TMP_DIR)
  saveRDS(x, f)
  xfun::embed_file(f, ...)
}

# embed plots
knit_error( {
  tmpfiles <- list()
  tmpfiles$ntprobnp <- tempfile(fileext = '.png', tmpdir = TMP_DIR)
  ggsave( plot = plots$ntprobnp, tmpfiles$ntprobnp, device = "png", bg = "white")
  embed_file(tmpfiles$ntprobnp, name = "NTproBNP.png", text = "[Get NTproBNP Distribution Plot as .png]")
}, 'Could not crate plot exports. Error: ')
# plot from cohort instead of plotlist ntprobnp
knit_error( {
  plots$ntprobnp_ecdf <- ecdf(cohort_report$NTproBNP.valueQuantity.value)
  embed_var(plots$ntprobnp_ecdf, name = "NTproBNP_ECDF_c.rds", text = "[Get NTproBNP Distribution Plot as .ecdf]")
}, 'Could not crate plot exports. Error: ')
knit_error( {
  tmpfiles$age <- tempfile(fileext = '.png', tmpdir = TMP_DIR)
  ggsave( plot = plots$age, tmpfiles$age, device = "png", bg = "white")
  embed_file(tmpfiles$age, name = "Age.png", text = "[Get Age Distribution Plot as .png]")
}, 'Could not crate plot exports. Error: ')
# plot from cohort instead of plotlist age
knit_error( {
  plots$age_ecdf <- ecdf(cohort_report$age)
  embed_var(plots$age_ecdf, name = "NTproBNP_ECDF.rds", text = "[Get Age Distribution Plot as .ecdf]")
}, 'Could not crate plot exports. Error: ')
knit_error( {
  embed_var(tables, name = "Table_list.rds", text = "[Get all Tables in List as .rds]")
}, 'Could not crate plot exports. Error: ')

```

# Extract [Pointblank](https://cran.r-project.org/web/packages/pointblank/index.html) Scans for Cohort and Diagnoses:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

# function that maps whitespace strings to NA then hashes all non NA values 
  digst_hash_without_NA <- function(x, ...) {
    if(is.na(x)) return(NA)
    x <- trimws(x)
    if(x == "") return(NA)
    return(digest(x, ...))
  }
  
  dir.create(OUTPUT_DIR_TMP)

knit_error({ 
  # pointblank data_scan
  # full cohort/diagnosis for integrity & completeness checks without metadata
  # fix typing cohort 
  cohort_full$NTproBNP.valueQuantity.value <- round(as.double(cohort_full$NTproBNP.valueQuantity.value), digits = -1)
  cohort_full$age <- as.integer(cohort_full$age)
  cohort_full$encounter.start <- parse_date_time(cohort_full$encounter.start, "ymd HMS", truncated = 5)
  cohort_full$encounter.end <- parse_date_time(cohort_full$encounter.end, "ymd HMS", truncated = 5)
  # hash ids before generating pointblank summary scan
  cohort_full$subject <- sapply(cohort_full$subject, digst_hash_without_NA, algo = "md5")
  cohort_full$encounter.id <- sapply(cohort_full$encounter.id, digst_hash_without_NA, algo = "md5")
  # save pointblank summary scan in temp file
  cohort_pointblank_scan <- pointblank::scan_data(cohort_full, sections = "OVM")
  cohort_pointblank_tmp <- tempfile(fileext = '.html', tmpdir = OUTPUT_DIR_TMP)
  pointblank::export_report(cohort_pointblank_scan, cohort_pointblank_tmp)
  embed_file(cohort_pointblank_tmp, name = "cohort_pointblank_scan.html", text = "[Get Pointblank scan for Cohort.csv as .html]")
}, 'Could not create Pointblank scan for Cohort.csv. Error: ')

knit_error({ 
  # diagnosis, typing all string is correct
  # hash ids before generating pointblank summary scan
  diagnoses_full$condition.id <- sapply(diagnoses_full$condition.id, digst_hash_without_NA, algo = "md5")
  diagnoses_full$subject <- sapply(diagnoses_full$subject, digst_hash_without_NA, algo = "md5")
  diagnoses_full$encounter.id <- sapply(diagnoses_full$encounter.id, digst_hash_without_NA, algo = "md5")
  # save pointblank summary scan in temp file
  diagnoses_pointblank_scan <- pointblank::scan_data(diagnoses_full, sections = "OVM")
  diagnoses_pointblank_tmp <- tempfile(fileext = '.html', tmpdir = OUTPUT_DIR_TMP)
  pointblank::export_report(diagnoses_pointblank_scan, diagnoses_pointblank_tmp)
  embed_file(diagnoses_pointblank_tmp, name = "diagnoses_pointblank_scan.html", text = "[Get Pointblank scan for Diagnoses.csv as .html]")
}, 'Could not create Pointblank scan for Diagnoses.csv. Error: ')

```

