---
title: "DQ report - SMITH"
output:
  html_document:
    df_print: paged
---


```{r include=FALSE}
### LIBRARIES ------------------------
library(knitr)
library(dataquieR)
library(data.table)
library(lubridate)

```

```{r include=FALSE}
### READ DATA -------------------

# only the fallback if we start knitr directly from R Studio
if (!exists("retrieve_file_cohort")) {
  retrieve_file_cohort <- paste0(OUTPUT_DIR_BASE, "/outputLocal/", PROJECT_NAME, "/Cohort.csv")
}

if(!file.exists(retrieve_file_cohort)){
  write("Ergebnisse nicht gefunden.", file = error_file)
  stop("Results not found - aborting.")
}

# read result csv
cohort_report <- read.csv(retrieve_file_cohort, sep = ";", colClasses = "character")

# read metadata
metadata_report <- read.csv("metadata.csv")
checks_report <- read.csv("contradictions.csv")
missings_report <- read.csv("missing-codes.csv")

# select columns supported with metadata from cohort
if(!all( metadata_report$VAR_NAMES %in% colnames(cohort_report) )){
  write("Metadaten passen nicht zur Kohorte.", file = error_file)
  stop("Metadata does not match Cohort - aborting.")
}
cohort_report <- cohort_report[,metadata_report$VAR_NAMES]

#extract known data types
datatyp_report <- as.data.table(t(metadata_report$DATA_TYPE))
colnames(datatyp_report) <- metadata_report$VAR_NAMES

#extract known value labels
value_labels <- as.data.table(t(metadata_report$VALUE_LABELS[2:7]))
colnames(value_labels) <- metadata_report$VAR_NAMES[2:7]
# function that returns only labels in string as list
get_labels <- function(string) {
  # regex to split at | num = : \s?\|?\s?\d\s\=\s
  # [-1] removes 1 = at start
  return(strsplit(string, "\\s?\\|?\\s?\\d\\s\\=\\s")[[1]][-1])
}
value_labels <- sapply(value_labels, get_labels)

#set up missing codes
missing_unknown <- missings_report$CODE_VALUE[2]
#item has integer/string type but cannot be converted to date for age calculation
missing_datetime <- missings_report$CODE_VALUE[3]
#item has correct type but is missing in value labels
missing_label <- missings_report$CODE_VALUE[4]


### Transform to dataquiR format -------------------

# function that sets integer values according to labels and missing codes
apply_labels <- function(vector, labels) {

  # unify casing, trim whitespace
  vector <- tolower(vector)
  labels <- tolower(labels)
  vector <- trimws(vector, which = "both")
  labels <- trimws(labels, which = "both")

  r <- NULL
  for(item in vector) {
    # default assumes value of item is missing in value labels
    value <- missing_label
    # if item is whitespace or NA set value to NA for correct dataquiR missingness
    if(item == "" || is.na(item) ) value <- NA
    else {
      i <- 0
      for(l in labels) {
        i <- i + 1
        # if item is found in labels set value to corresponding integer
        if(item == l) value <- i
      }
    }
    r <- c(r,value)
  }

  return(r)
}

# age not calculate from birthdate anymore
# calculate_age <- function(vector) {
#
#  # ws is assinged missing code for wrong format as well
#  # # unify casing, trim whitespace
#  # vector <- tolower(vector)
#  # vector <- trimws(vector, which = "both")
#
#  #convert to datetime with at least years (YYYY or any YYYY-MM... format)
#  r <- parse_date_time(vector, "ymd", truncated = 2)
#
#  # calculate age with lubridate period
#  r <- lapply(r, function(x) {as.period(interval(x, Sys.Date()), unit = "year")$year} )
#  r <- unlist(r)
#
#  for (i in 1:length(r) ) {
#    if( is.na( r[i] ) && !is.na(vector[i]) ) r[i] <- missing_datetime
#  }
#
#  return(r)
#}
#
# table(cohort_report$NTproBNP.unit)

#apply labels and fix typing per column - missmatched typing is returned as NA (NOT as missing code)
#cohort_report$NTproBNP.date <- parse_date_time(cohort_report$NTproBNP.date, "ymd HMS", truncated = 5)
cohort_report$NTproBNP.valueQuantity.value <- as.double(cohort_report$NTproBNP.valueQuantity.value)
cohort_report$NTproBNP.code <- as.integer( apply_labels(cohort_report$NTproBNP.code, value_labels$NTproBNP.code) )
cohort_report$NTproBNP.codeSystem <- as.integer(  apply_labels(cohort_report$NTproBNP.codeSystem, value_labels$NTproBNP.codeSystem) )
cohort_report$NTproBNP.unit <- as.integer( apply_labels(cohort_report$NTproBNP.unit, value_labels$NTproBNP.unit) )
cohort_report$NTproBNP.unitSystem <- as.integer( apply_labels(cohort_report$NTproBNP.unitSystem, value_labels$NTproBNP.unitSystem) )
cohort_report$gender <- as.integer( apply_labels(cohort_report$gender, value_labels$gender) )
cohort_report$age <- as.integer(cohort_report$age)
cohort_report$encounter.start <- parse_date_time(cohort_report$encounter.start, "ymd HMS", truncated = 5)
cohort_report$encounter.end <- parse_date_time(cohort_report$encounter.end, "ymd HMS", truncated = 5)
cohort_report$NTproBNP.valueQuantity.comparator <- as.integer( apply_labels(cohort_report$NTproBNP.valueQuantity.comparator, value_labels$NTproBNP.valueQuantity.comparator) )

# verify correct typing
#sapply(cohort_report, class)

```

```{r include=FALSE}
### GENERATE REPORTS ----------------------

#function for logging errors inside knit output
knit_error <- function(f ,er_message) {
   tryCatch(
      { f },
        error = function(e){
            print(er_message)
            print(e)
        }
    )
}

knit_error( smith_dq_report <- dq_report(study_data = cohort_report,
                          meta_data  = metadata_report,
                          check_table = checks_report,
                          MISSING_LIST = missings_report,
                          label_col  = LABEL) ,
'Could not generate DQ report for Cohort. Error: ')



```

## Data Quality report for the 6. Projectathon of MII: SMITH

#### This report was created with the [dataquieR](https://dataquality.ship-med.uni-greifswald.de/tutorials2.html) Package for R.

## Result of contradiction checks:
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error( smith_dq_report$long_format$con_contradictions$results[[1]]$SummaryPlot, 'Could not print Contradictions SummaryPlot for Cohort. Error: ')


```

### Contradiction checks as Table:

```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error( smith_dq_report$long_format$con_contradictions$results[[1]]$SummaryTable, 'Could not print Contradictions SummaryTable for Cohort. Error: ')

```

## Results of limit checks:
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error(con_limit_deviations(study_data = cohort_report, meta_data = metadata_report, resp_vars = 'NTproBNP.valueQuantity.value')$SummaryPlotList$NTproBNP.valueQuantity.value, 'Could not print Limit SummaryPlot for Cohort. Error: ')

```

### Limit checks as Table:

```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(con_limit_deviations(study_data = cohort_report, meta_data = metadata_report, resp_vars = 'NTproBNP.valueQuantity.value')$SummaryTable, 'Could not print Limit SummaryPlot for Cohort. Error: ')

```

### Outlier Detection NTproBNP:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(acc_univariate_outlier(resp_vars = 'NTproBNP.valueQuantity.value',
                         study_data = cohort_report,
                         meta_data = metadata_report
                       )$SummaryPlotList$NTproBNP.valueQuantity.value, 'Could not print NTproBNP Outlier SummaryPlot for Cohort. Error: ')

```

### Outlier Detection NTproBNP as Table:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(acc_univariate_outlier(resp_vars = 'NTproBNP.valueQuantity.value',
                         study_data = cohort_report,
                         meta_data = metadata_report
                       )$SummaryTable, 'Could not print NTproBNP Outlier SummaryTable for Cohort. Error: ')

```

# Information about distribution:

## Age:
```{r message=FALSE, warning=FALSE,echo=FALSE}
knit_error(acc_plots <- acc_distributions(resp_vars      = NULL, 
                         label_col      = "LABEL",
                         study_data     = cohort_report, 
                         meta_data      = metadata_report), 
           'Could not generate SummaryPlotList$gender for Cohort. Error: ')


knit_error(acc_plots$SummaryPlotList$age, 'Could not print SummaryPlotList$age for Cohort. Eorror: ')

```

### Outlier Detection Age:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(acc_univariate_outlier(resp_vars = 'age',
                         study_data = cohort_report,
                         meta_data = metadata_report
                       )$SummaryPlotList$age, 'Could not print Age Outlier SummaryPlot for Cohort. Error: ')

```

### Outlier Detection Age as Table:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(acc_univariate_outlier(resp_vars = 'age',
                         study_data = cohort_report,
                         meta_data = metadata_report
                       )$SummaryTable, 'Could not print Age Outlier SummaryTable for Cohort. Error: ')

```


## Gender:
#### 1 = male | 2 = female | 3 = other | 4 = unknown
```{r message=FALSE, warning=FALSE,echo=FALSE}

knit_error(acc_plots <- acc_distributions(resp_vars      = NULL, 
                         label_col      = "LABEL",
                         study_data     = cohort_report, 
                         meta_data      = metadata_report), 
           'Could not generate SummaryPlotList$gender for Cohort. Error: ')

print('Gender as table:')
table(smith_dq_report$study_data$gender)

knit_error(acc_plots$SummaryPlotList$gender, 'Could not print SummaryPlotList$gender for Cohort. Error: ')

```

## LOINC Codes:
#### 1 = other | 2 = 33763-4 | 3 = 71425-3 | 4 = 33762-6 | 5 = 83107-3 | 6 =  83108-1 | 7 = 77622-9 | 8 = 77621-1
```{r message=FALSE, warning=FALSE,echo=FALSE}

print('LOIC Codes as table:')
knit_error(table(smith_dq_report$study_data$NTproBNP.code), 'Could not tab SummaryPlotList$NTproBNP.code for Cohrot. Error: ')

print('If less then two Codes are used no plot will be generated.')
knit_error(acc_plots$SummaryPlotList$NTproBNP.code, 'Could not print SummaryPlotList$NTproBNP.code for Cohrot. Error: ')

```

## Comperators:
#### 1 = other | 2 = < | 3 = <= | 4 = > | 5 = >=
```{r message=FALSE, warning=FALSE,echo=FALSE}

print('Comperators as table:')
knit_error(table(smith_dq_report$study_data$NTproBNP.valueQuantity.comparator), 'Could not tab SummaryPlotList$study_data$NTproBNP.valueQuantity.comparator for Cohrot. Error: ')

print('If less then two Comperaors are used no plot will be generated.')
knit_error(acc_plots$SummaryPlotList$NTproBNP.valueQuantity.comparator, 'Could not print SummaryPlotList$comparator for Cohort. Error: ')

```

## Units:
#### 1 = other | 2 = pg/mL | 3 = pg/dL | 4 = pg/100mL | 5 = pg/L | 6 = pg% | 7 = pmol/L | 8 = ng/L | 9 = Mt/m3
```{r message=FALSE, warning=FALSE,echo=FALSE}

print('Units as table:')
knit_error(table(smith_dq_report$study_data$NTproBNP.unit), 'Could not tab SummaryPlotList$study_data$NTproBNP.unit for Cohrot. Error: ')

print('If less then two Units are used no plot will be generated.')
knit_error(acc_plots$SummaryPlotList$NTproBNP.unit, 'Could not print SummaryPlotList$NTproBNP.unit for Cohort. Error: ')

```

# Information about completness:

```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

#Table
knit_error(smith_dq_report$long_format$com_item_missingness$results[[1]]$SummaryTable, 'Could not print Missingness Table for Cohort. Error: ')


```

# Completness with missing codes:
#### MISSING 9991100: Missing - reason unknown
#### MISSING 9991200: Deleted - Datetime format not supported
#### MISSING 9991300: Deleted - not in VALUE_LABELS
Regarding completeness and missing codes:
If a field has the wrong datatype it is evaluated to NA and tracked under Sysmiss in the completeness table. If a field has the correct typing but an undefined value (for example any string as a gender other than 'male', 'female', 'other', 'unknown') the missing code 9991300 is used instead.
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

#Plot
knit_error(smith_dq_report$long_format$com_item_missingness$results[[1]]$ReportSummaryTable, 'Could not print Missingness SummaryPlot for Cohort. Error: ')


```

# Further possible analysis with dataquieR:
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

knit_error(smith_dq_report$app_mat$ApplicabilityPlot, 'Could not print ApplicabilityPlot for Cohort. Error: ')

```
